### 2024-02-20 paper_sisc_first_revision/localpsf.tex
not localized -> widely dispersed
a H-matrix -> an H-matrix

It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region. In principle, the PSF-based method can be applied even if the impulse responses are not localized. However, in this case only a small number of impulse responses can be computed per batch, which means more batches, and hence more operator applies, are needed to form an accurate approximation.

@@

The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form,~\eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
$\Aker_h$, which we do not know, and which differs from $\Aker$ due to discretization error. Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^\fedim_{\massmatrix^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,\fedim$. Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
$
\boldsymbol{\rho}^* = \massmatrix^{-1} \boldsymbol{\rho}.
$
Therefore, the formula for the impulse response from~\eqref{eq:impulse_response_delta_action} becomes 
$
\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
$
and the $(y,x)$ kernel entry of $\Aker_h$ may be written as
$
	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
$
Now define $\mathbf{\Aker} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Aker}_{ij} := \Aker_h(\fepoint_i, \fepoint_j).
\end{equation}
Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{\fepoint_i} = \mathbf{e}_i$. Thus, we have
$
	\Aker_h(\fepoint_i,\fepoint_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
$
which implies
\begin{equation}
\label{eq:boldA}
	\mathbf{A} = \massmatrix \mathbf{\Aker} \massmatrix.
\end{equation}
Broadly, we will construct an H-matrix approximation of $\mathbf{A}$ by forming a H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods. Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible. The bulk of the PSF-based method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.

@@

We form an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming an H-matrix representation $\mathbf{\Aker}_H$ of $\mathbf{\Aker}$ then multiplying $\mathbf{\Aker}$ with mass matrices $\massmatrix$ per~\eqref{eq:boldA} to form
$
\mathbf{A}_H = \massmatrix \mathbf{\Aker}_H \massmatrix.
$
Here, we use a diagonal lumped mass matrix, so these matrix-matrix multiplications are trivial. If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
We use H1 matrices in the numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.

@@

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\mathbf{H}_\text{gn}$ that arises from the data misfit). The 
approximations are indicated by ``PSF ($\nbatch$)'', where $\nbatch$ is the number of impulse response batches used to build the approximation. The Hessian of the regularization term is a combination of
stiffness and mass matrices, which are sparse. Therefore, we form
H-matrix representations of these matrices and combine them into a H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
standard sparse H-matrix techniques. Then, H-matrix approximations of the Gauss-Newton Hessian, 
$\mathbf{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H$,
are formed by adding $\mathbf{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic. We modify $\widetilde{\mathbf{H}}$ to be (approximately) symmetric positive semi-definite via the procedure described in Section~\ref{sec:flip_eigs}. We factor $\mathbf{\preconditioner}$
using fast H-matrix methods, then use the factorization as a
preconditioner. We approximate $\mathbf{H}_\text{gn}$ rather than $\mathbf{H}$ because $\mathbf{H}$ more often has negative values in its integral kernel. The numerical results show that~$\widetilde{\mathbf{H}}$ is a good preconditioner for both~$\mathbf{H}_\text{gn}$ and~$\mathbf{H}$.

>>>>

It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region. In principle, the PSF-based method can be applied even if the impulse responses are widely dispersed. However, in this case only a small number of impulse responses can be computed per batch, which means more batches, and hence more operator applies, are needed to form an accurate approximation.

@@

The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form,~\eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
$\Aker_h$, which we do not know, and which differs from $\Aker$ due to discretization error. Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^\fedim_{\massmatrix^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,\fedim$. Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
$
\boldsymbol{\rho}^* = \massmatrix^{-1} \boldsymbol{\rho}.
$
Therefore, the formula for the impulse response from~\eqref{eq:impulse_response_delta_action} becomes 
$
\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
$
and the $(y,x)$ kernel entry of $\Aker_h$ may be written as
$
	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
$
Now define $\mathbf{\Aker} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Aker}_{ij} := \Aker_h(\fepoint_i, \fepoint_j).
\end{equation}
Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{\fepoint_i} = \mathbf{e}_i$. Thus, we have
$
	\Aker_h(\fepoint_i,\fepoint_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
$
which implies
\begin{equation}
\label{eq:boldA}
	\mathbf{A} = \massmatrix \mathbf{\Aker} \massmatrix.
\end{equation}
Broadly, we will construct an H-matrix approximation of $\mathbf{A}$ by forming an H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods. Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible. The bulk of the PSF-based method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.

@@

We form an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming an H-matrix representation $\mathbf{\Aker}_H$ of $\mathbf{\Aker}$ then multiplying $\mathbf{\Aker}$ with mass matrices $\massmatrix$ per~\eqref{eq:boldA} to form
$
\mathbf{A}_H = \massmatrix \mathbf{\Aker}_H \massmatrix.
$
Here, we use a diagonal lumped mass matrix, so these matrix-matrix multiplications are trivial. If a non-diagonal mass matrix is used, one may form an H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
We use H1 matrices in the numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.

@@

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\mathbf{H}_\text{gn}$ that arises from the data misfit). The 
approximations are indicated by ``PSF ($\nbatch$)'', where $\nbatch$ is the number of impulse response batches used to build the approximation. The Hessian of the regularization term is a combination of
stiffness and mass matrices, which are sparse. Therefore, we form
H-matrix representations of these matrices and combine them into an H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
standard sparse H-matrix techniques. Then, H-matrix approximations of the Gauss-Newton Hessian, 
$\mathbf{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H$,
are formed by adding $\mathbf{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic. We modify $\widetilde{\mathbf{H}}$ to be (approximately) symmetric positive semi-definite via the procedure described in Section~\ref{sec:flip_eigs}. We factor $\mathbf{\preconditioner}$
using fast H-matrix methods, then use the factorization as a
preconditioner. We approximate $\mathbf{H}_\text{gn}$ rather than $\mathbf{H}$ because $\mathbf{H}$ more often has negative values in its integral kernel. The numerical results show that~$\widetilde{\mathbf{H}}$ is a good preconditioner for both~$\mathbf{H}_\text{gn}$ and~$\mathbf{H}$.
